---
title: "R Notebook"
output: html_notebook
---
#Importing tables
Metadata and resistome
```{r}
library(readxl)
library(tibble)
library(dplyr)
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/tables/metagenome_shotgun")
metadata <- read_excel("contigs_iguanas_tables.xlsx", sheet = "metadata") 
deepARG <- read_excel("contigs_iguanas_tables.xlsx", sheet = "deepARG")
```

#OTU TABLE raw reads
Resistome abundance
```{r}
otu_table <- read_excel("contigs_iguanas_tables.xlsx", sheet = "ARG_abundance") %>% tibble::remove_rownames() %>%  tibble::column_to_rownames("deepARG_best-hit")
otu_table[is.na(otu_table)] = 0
otu_table=as.matrix(otu_table,rownames=TRUE)
```

#OTU TABLE normalized by the number of bacterial reads
```{r}
norm_abundance <- merge(deepARG, metadata, by = "sample_id") %>% 
  select(sample_id, `deepARG_best-hit`, deepARG_counts, Bacteria_bracken) %>% 
  group_by(sample_id, `deepARG_best-hit`) %>% 
  summarise(norm.abundance = sum(deepARG_counts/Bacteria_bracken)) %>% 
  as.data.frame()

library(tidyr)
deepARG.norm.mx <-spread(norm_abundance, key = sample_id, value = norm.abundance) %>%  tibble::remove_rownames() %>%  tibble::column_to_rownames("deepARG_best-hit")
  
deepARG.norm.mx[is.na(deepARG.norm.mx)] = 0
deepARG.norm.mx=as.matrix(deepARG.norm.mx,rownames=T)
deepARG.norm.mx <- deepARG.norm.mx*10^32 #To convert numbers to integers
```


#Calculuating normalize mean abundance per location
```{r}
merge(deepARG, metadata, by = "sample_id") %>% 
  select(sample_id, `deepARG_best-hit`, Location, deepARG_counts, Bacteria_bracken) %>%
  group_by(sample_id, Location) %>% 
  summarise(norm.abundance = sum(deepARG_counts/Bacteria_bracken*100)) %>% 
  as.data.frame() %>% 
  group_by(Location) %>% 
  summarise(mean_ab = (norm.abundance), sd =sd(norm.abundance))
```


#Plot of normalize mean abundance per location
```{r}
my_comparisons <- list(c("Cabo Douglas", "Punta Espinoza"), c("Punta Espinoza", "Rabida"), c("Cabo Douglas", "Rabida"))

norm_ab_arg_plot <- merge(deepARG, metadata, by = "sample_id") %>% 
  select(sample_id, `deepARG_best-hit`, Location, deepARG_counts, Bacteria_bracken) %>%
  group_by(sample_id, Location) %>% 
  summarise(norm.abundance = sum(deepARG_counts/Bacteria_bracken*100)) %>% 
    mutate(Type = "ARGs") %>% 
  as.data.frame() %>% 
  ggboxplot(x = "Location", y = "norm.abundance", add = "jitter", palette = "aaas", fill = "Location", color  = "Location", ylab = "Normalized abundance %", alpha = 0.5, xlab = F, facet.by = "Type") +
  stat_compare_means(comparisons = my_comparisons, 
                     tip.length=.02, hide.ns = F, size = 3.5) +
  stat_compare_means(label.y.npc = "bottom", label.x.npc = .3)
norm_ab_arg_plot
```

#Saving resistome abundance plot
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/figures-plots/")
ggsave("deepARG_norm.abundance.png", plot = norm_ab_plot, width = 6, height = 4)
```

#Making taxonomy table
```{r}
taxonomy <- deepARG %>% select(`deepARG_best-hit`,deepARG_Class,deepARG_ARG) %>% 
  distinct() %>% #Remove duplicate rows
  arrange(`deepARG_best-hit`) %>% #organize names alphabetically
  rename(OTU = `deepARG_best-hit`, Class = deepARG_Class, ARG = deepARG_ARG) %>% #Rename columns
  tibble::remove_rownames() %>%  tibble::column_to_rownames("OTU")
taxonomy=as.matrix(taxonomy,rownames=TRUE)
```

#Phyloseq object = Resistome
```{r}
library(phyloseq)
#import as phyloseq objects
OTU = otu_table(deepARG.norm.mx,taxa_are_rows=T)
TAX = tax_table(taxonomy)
META = metadata %>% tibble::remove_rownames() %>% tibble::column_to_rownames("sample_id") %>% sample_data()

physeq=phyloseq(OTU,TAX,META)
```

#Alpha diversity
```{r fig.width=10, fig.height=40}
alpha_diversity <- estimate_richness(physeq, measures = c("Shannon", "Observed", "Chao1"))
df_alpha <- data.frame(alpha_diversity, sample_data(physeq))
alpha_table <- reshape2::melt(df_alpha, measure.var=c("Shannon","Observed","Chao1"),id.vars=c("Location","Island")) %>% 
  rename(Index = variable)
alpha_table$value = as.numeric(alpha_table$value)
```

#Alpha diversity plot
```{r fig.height=5, fig.width=14}
library(ggplot2)
library(ggpubr)
library(ggsci)

my_comparisons <- list(c("Cabo Douglas", "Punta Espinoza"), c("Punta Espinoza", "Rabida"), c("Cabo Douglas", "Rabida"))

##################### Running box plot #############################
alpha_plot <- alpha_table %>% #shannon_box, chao1_box
  filter(Index %in% c("Shannon","Chao1")) %>% 
  ggboxplot(x = "Location", y = "value", fill = "Location", add = "jitter", 
            alpha=0.5, color = "Location", ylab = "Value", 
            facet.by = "Index", scales ="free_y") +
  theme(legend.position = "none",axis.title.x=element_blank()) + 
  stat_compare_means(comparisons = my_comparisons, 
                     tip.length=.02, hide.ns = F) +
  stat_compare_means(label.y.npc = "bottom", label.x.npc = .3) +
   scale_fill_aaas() + scale_color_aaas()
alpha_plot
```


#Bray-Curtis dissimilarity statistics
```{r}
## PERMANOVA
library(vegan)
bray_dist = phyloseq::distance(physeq, method="bray")
adonis_bray = adonis(bray_dist ~ sample_data(physeq)$Location)
adonis_bray

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
bray_disp <-betadisper(bray_dist, sample_data(physeq)$Location, type=c("median"))
anova(bray_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(bray_disp, label = T, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(bray_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(bray_disp)

scores(bray_disp, 1:4, display = "centroids")
```

##Bray-Curtis dissimilarity (beta diversity) plot
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.12, ", paste(italic('p')),"=0.001"))

beta_bray <- ordinate(physeq, "PCoA", "bray") #RUN this only ONCE because it takes a lot of time
beta_bray_plot = plot_ordination(physeq, beta_bray, type="Location", color="Location") + 
  stat_ellipse(type = "t", linetype = 1) + 
  theme_classic()+
  labs(colour = "Location") +#To add arrows https://neavemj.github.io/posts/coralMicrobiome
  annotate("text", x = 0, y = 0.15, label = label_perm, colour = "black") +
  scale_color_aaas()

beta_bray_plot
```

#Arranging alpha and beta diversity plots into one
```{r fig.width=9, fig.height=5}
diversity = ggarrange(alpha_plot, beta_bray_plot, nrow = 2, ncol = 1, labels = c("A","B"))
diversity
```

#Saving plots as PNG
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/figures-plots/")
ggsave("deepARG_assembly_diversity_norm.bacteria.png", plot = diversity, width = 9, height = 8)
```

#ANCOM-BC
```{r}
library(microbiome)
library(nloptr)
library(ANCOMBC)
```

#Filtering samples and taxa for further analyses
```{r}
`%notin%` <- Negate(`%in%`)
physeq_Type = phyloseq::tax_glom(physeq, taxrank = "Type")
physeq_Gene = phyloseq::tax_glom(physeq, taxrank = "Gene")
```

#Type
```{r}
physeq <- physeq_Type
`%notin%` <- Negate(`%in%`)
physeq_CD_RA = subset_samples(physeq, Location %notin%"Punta Espinoza")
physeq_CD_PE = subset_samples(physeq, Location %notin% "Rabida")
physeq_PE_RA = subset_samples(physeq, Location %notin% "Cabo Douglas")
```


#Fernandina vs. Rabida
```{r}
out = ancombc(phyloseq = physeq, formula = "Island",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Island", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
#write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_Fernandina_vs_Rabida_TYPE.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Cabo Douglas vs. Rabida
```{r}
physeq <- physeq_CD_RA

out = ancombc(phyloseq = physeq, formula = "Location",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Location", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_CD_vs_RA_TYPE.csv",sep = ",", row.names = F)
```


#Cabo Douglas vs. Punta Espinoza
```{r}
physeq <- physeq_CD_PE

out = ancombc(phyloseq = physeq, formula = "Location",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Location", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
#write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_CD_vs_PE_TYPE.csv",sep = ",", row.names = F)
```


#Rabida vs. Punta Espinoza
```{r}
physeq <- physeq_PE_RA

out = ancombc(phyloseq = physeq, formula = "Location",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Location", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
#write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_RA_vs_PE_TYPE.csv",sep = ",", row.names = F)
```

#Gene
```{r}
physeq <- physeq_Gene
`%notin%` <- Negate(`%in%`)
physeq_CD_RA = subset_samples(physeq, Location %notin%"Punta Espinoza")
physeq_CD_PE = subset_samples(physeq, Location %notin% "Rabida")
physeq_PE_RA = subset_samples(physeq, Location %notin% "Cabo Douglas")
```


#Fernandina vs. Rabida
```{r}
out = ancombc(phyloseq = physeq, formula = "Island",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Island", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_Fernandina_vs_Rabida_GENE.csv",sep = ",", row.names = F) #Writing `sigtab` to my tables
```

#Cabo Douglas vs. Rabida
```{r}
physeq <- physeq_CD_RA

out = ancombc(phyloseq = physeq, formula = "Location",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Location", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_CD_vs_RA_GENE.csv",sep = ",", row.names = F)
```


#Cabo Douglas vs. Punta Espinoza
```{r}
physeq <- physeq_CD_PE

out = ancombc(phyloseq = physeq, formula = "Location",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Location", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_CD_vs_PE_GENE.csv",sep = ",", row.names = F)
```


#Rabida vs. Punta Espinoza
```{r}
physeq <- physeq_PE_RA

out = ancombc(phyloseq = physeq, formula = "Location",
              p_adj_method = "holm", zero_cut = 0.90, lib_cut = 0,
              group = "Location", struc_zero = F, neg_lb = FALSE,
              tol = 1e-05, max_iter = 100, conserve = TRUE,
              alpha = 0.05, global = F)
res = out$res

ancom_results = res %>% 
  as_tibble()

ancom_signif_list <- ancom_results %>% 
  filter(p_val <= 0.05)

ancom_signif_names <- data.frame(sapply(ancom_signif_list,c)) 
row.names(ancom_signif_names) <- rownames(ancom_signif_list$beta)
ancom_signif_names <- merge(ancom_signif_names,physeq@tax_table, by = "row.names")
write.table(ancom_signif_names, file="~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_iguana/tables/metagenome_shotgun/MGEs/ANCOMBC_RA_vs_PE_GENE.csv",sep = ",", row.names = F)
```

#COMPOSITION
##Class data frame

```{r}
physeq_Class = phyloseq::tax_glom(physeq, taxrank = "Class")
data_Class <- psmelt(physeq_Class) # create dataframe from phyloseq object

library(dplyr)
data_Class_rel = data_Class %>% group_by(Sample) %>% mutate(relAbundBySample = Abundance / sum(Abundance)*100) #Converting counts to relative abundance %
data_Class_rel = data_Class_rel %>% mutate(Abundance = Abundance / 10^32) #Converting counts to normalized abundance since I multiplied the normalized abundance by 10^9 to calculate the Shannon indes
```

```{r fig.width=12}
Class_ab <- ggbarplot(data=data_Class_rel, x= "Sample", y = "relAbundBySample", color = "Class", fill = "Class", palette = get_palette("simpsons",22), ylab = "Abundance", 
                          #order = metadata$Sample_ID
                          ) + 
  facet_grid(~Location, scales = "free_x", space = "free")+labs(colour = "Class", fill = "Class") + 
  theme(axis.text.x = element_text(angle = 90, size = 7, hjust = 1, colour = "gray35"))

Class_ab
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/figures-plots/")
ggsave(plot = Class_ab, "deepARG_Class_relab.png", width = 12, height = 6)
```

```{r}
data_ARG <- psmelt(physeq_ARG) # create dataframe from phyloseq object

library(dplyr)
data_ARG_rel = data_ARG %>% group_by(Sample) %>% mutate(relAbundBySample = Abundance / sum(Abundance)*100) #Converting counts to relative abundance %
data_ARG_rel = data_ARG_rel %>% mutate(Abundance = Abundance / 10^32) #Converting counts to normalized abundance since I multiplied the normalized abundance by 10^9 to calculate the Shannon indes
```

```{r fig.width=12}
data_ARG_count  <- data_ARG_rel %>% 
  group_by(`ARG`) %>%
  summarize(sum_genes=sum(Abundance))

top20 <- top_n(data_ARG_count,20)

ARG_barplot <- data_ARG_rel %>% 
  filter(`ARG` %in% top20$`ARG`) %>% 
  ggbarplot(data=, x= "Sample", y = "relAbundBySample", color = "ARG", fill = "ARG", palette = get_palette("simpsons",22), ylab = "Abundance", legend = "right"
                          #order = metadata$Sample_ID
                          ) + 
  facet_grid(~Location, scales = "free_x", space = "free")+labs(colour = "ARG", fill = "ARG") + 
  theme(axis.text.x = element_text(angle = 90, size = 7, hjust = 1, colour = "gray35"))
ARG_barplot
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/figures-plots/")
ggsave(plot = ARG_barplot, "deepARG_Gene_relab.png", width = 12, height = 6)
```

#HEAT TREE
##Modifying phyloseq object
```{r}
physeq_t  = physeq_ARG %>% transform_sample_counts(function(x) x / sum(x)/32*100 ) 
```

##Metacoder
```{r}
library(metacoder)

#transforming physeq object into a metacoder object

metacoder_object = parse_phyloseq(physeq_t) 

#creating a column with total counts. Keep as is
metacoder_object$data$taxon_counts <- calc_taxon_abund(metacoder_object, data = "otu_table") 

metacoder_object$data$taxon_counts$total <- rowSums(metacoder_object$data$taxon_counts[, -1]) # -1 = taxon_id column

metacoder_object$data$taxon_prop <- calc_obs_props(metacoder_object, "otu_table")
```
##Heat tree
```{r}
#Plotting heat-tree

heat_tree = metacoder_object %>%
  # subset to the class rank
  heat_tree(node_label = taxon_names,
            node_size = metacoder_object$data$taxon_counts$total,
         #   node_size_range = c(0.01, 0.5),
          #  edge_size_range = c(0.001, 0.005),
            node_label_size_range = c(0.0001, 0.05),
         node_color = metacoder_object$data$taxon_counts$total,

            node_color_axis_label = "Proportion",
            node_size_axis_label = "Taxa count",
            overlap_avoidance =1) #Modify if necssary
heat_tree
```

```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/figures-plots/") 

library(ggplot2)
ggsave(plot = heat_tree, "heat-tree_ARGs.png", height = 9, width = 11)
```

