---
title: "R Notebook"
output: html_notebook
---
#Libraries needed
```{r}
library("phyloseq")
library("ggplot2")
library("ape")
library("gridExtra")
library("ggpubr")
library("ggsignif")
library(forcats)
library(vegan)
library(cowplot)
```

#Phyloseq object
```{r}
setwd("~/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/tables/microbiome_16S/qiime2_phyloseq")

#read in otu table
otu_table = read.table("feature-table.txt",sep="\t",row.names = 1,header = TRUE)
otu_table=as.matrix(otu_table,rownames=TRUE)

#read in taxonomy
taxonomy = read.csv("taxonomy.tsv",sep="\t",row.names=1)
taxonomy=as.matrix(taxonomy)

#read in metadata
metadata=read.table("metadata_marine-iguana_2021-04-08.csv", sep = ",", header = TRUE, row.names=1, na.strings = "") 
metadata
phy_tree = read_tree("tree.nwk")

#import as phyloseq objects
OTU = otu_table(otu_table,taxa_are_rows=TRUE)
TAX = tax_table(taxonomy)
META = sample_data(metadata)

#Physeq object
physeq=phyloseq(OTU,TAX,META,phy_tree)
```

#Fitering Chloroplast and Mitocondria OTUs
```{r}
`%notin%` <- Negate(`%in%`)

#Filtering samples
physeq_filtered = subset_taxa(physeq, Order %notin% "Chloroplast", Family %notin% "Mitochondria")
physeq = physeq_filtered
```

#Alpha diversity
```{r fig.width=10, fig.height=40}
alpha_diversity <- estimate_richness(physeq, measures = c("Shannon", "Chao1"))
df_alpha <- data.frame(alpha_diversity, sample_data(physeq))
df <- reshape2::melt(df_alpha, measure.var=c("Shannon","Chao1"),id.vars=c("location","island"))
df$value = as.numeric(df$value)

#selecting the rows corresponding to each index from the main dataframe
shannon = dplyr::filter(df, variable == "Shannon")
chao = dplyr::filter(df, variable == "Chao1")
```


##Alpha diversity gut plot
```{r}
library(ggsci)

my_comparisons <- list(c("Cabo Douglas", "Punta Espinoza"), c("Punta Espinoza", "Rabida"), c("Cabo Douglas", "Rabida"))

##################### Running box plot #############################

shannon_plot <- shannon %>% #shannon_box, chao1_box
  ggplot(aes(x = location, y = value, fill = location)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(color="black", size=1, alpha=0.5) +
  labs(y="Shannon", x="location") +
  annotate("text", x = 2, y = 3.2, label = "Kruskal-Wallis, p = 0.0097", colour = "black") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme_classic() + 
  theme(legend.position = "none",axis.title.x=element_blank()) + 
  stat_compare_means(comparisons = my_comparisons, label.y = c(2.95,3,3.05), 
                     tip.length=.02, hide.ns = F, size =3) +
   scale_fill_aaas()
shannon_plot

chao_plot <- chao %>% #shannon_box, chao1_box
  ggplot(aes(x = location, y = value, fill = location)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(color="black", size=1, alpha=0.5) +
  labs(y="Chao1", x="location") +
  annotate("text", x = 2, y = 130, label = "Kruskal-Wallis, p = 0.0032", colour = "black") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme_classic() + 
  theme(legend.position = "none",axis.title.x=element_blank()) + 
  stat_compare_means(comparisons = my_comparisons, label.y = c(105,110,115), 
                     tip.length=.02, hide.ns = F, size =3) +
   scale_fill_aaas()
 chao_plot
```

##Arranging both diversity plots
```{r}
alpha_div_plots = ggarrange(shannon_plot,chao_plot,nrow = 2, ncol = 1, align="hv",labels = c("A","B")) 
alpha_div_plots
```

#Beta Diversity

##Weighted Unifrac stats 
Analyzed by location
```{r}
## PERMANOVA
library(vegan)
wuinfrac_dist = phyloseq::distance(physeq, method="wunifrac") #RUN this only once because it takes a lot of time
adonis_wunifrac = adonis(wuinfrac_dist ~ sample_data(physeq)$location)
adonis_wunifrac

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
wuni_disp <-betadisper(wuinfrac_dist, sample_data(physeq)$location, type=c("median"))
anova(wuni_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(wuni_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(wuni_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(wuni_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(wuni_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(wuni_disp)

scores(wuni_disp, 1:4, display = "centroids")

rda(otu_table)
```

##Weighted unifrac beta diversity plot with phyloseq
Used to check the PCoA %
```{r}
beta_wu <- ordinate(physeq, "PCoA", "wunifrac")
beta_wu_plot = plot_ordination(physeq, beta_wu, type="location", color="location", shape="location", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "location") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_wu_plot
```

#Final weighted unifrac plot
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.16, ", paste(italic('p')),"=0.001"))

beta_scatter = as.data.frame(beta_wu[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
pmain_wuF = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "location", palette = "aaas",ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +labs(x = "PCoA 1 (40.2%) ", y = "PCoA 2 (19.3%)", colour = "location", fill = "location") +annotate("text", x = -0.04, y = -0.06, label = label_perm, colour = "black")
pmain_wuF
```

##Bray-Curtis dissimilarity statistics
```{r}
## PERMANOVA
library(vegan)
bray_dist = phyloseq::distance(physeq, method="bray") #RUN this only once because it takes a lot of time
adonis_bray = adonis(bray_dist ~ sample_data(physeq)$location)
adonis_bray

## Significant PERMANOVA indicates that centroid (or spatial median) among groups is different and/or with-group dispersion among groups is different

## PERMDISP
bray_disp <-betadisper(wuinfrac_dist, sample_data(physeq)$location, type=c("median"))
anova(bray_disp)

## If PERMANOVA and PERMDISP are both significant, you can use plotting to tell if PERMANOVA was significant based on centroid (or spatial median)
plot(bray_disp)
#?plot.betadisper

## Would look better with higher replication for groups
plot(bray_disp, label = F)

## Plot with 1 standard deviation ellipses around the group medians
## sample size issue here, but you get the idea
plot(bray_disp, label = F, hull = F, ellipse = T)

## Within-group dispersion that PERMDISP is testing
boxplot(bray_disp, las = 2, cex.lab=1.5)
?boxplot

## pairwise p-values
TukeyHSD(bray_disp)

scores(bray_disp, 1:4, display = "centroids")
```

##Bray beta diversity plot
```{r}
beta_bray <- ordinate(physeq, "PCoA", "bray") #RUN this only ONCE because it takes a lot of time
beta_bray_plot = plot_ordination(physeq, beta_bray, type="location", color="location", shape="location", title="PCoA Weighted Unifrac") + stat_ellipse(type = "t", linetype = 3) + stat_ellipse(type = "t") + theme_bw()+labs(colour = "location") #To add arrows https://neavemj.github.io/posts/coralMicrobiome
beta_bray_plot
```

##Bray-Curtis final plot
```{r}
label_perm <- expression(paste("PERMANOVA, ",R^2 ,"= 0.12, ", paste(italic('p')),"=0.001"))
#library(lemon)

beta_scatter = as.data.frame(beta_bray[["vectors"]])
beta_meta = merge(beta_scatter,metadata,by = 0,all=F)
pmain_brayF = ggscatter(beta_meta, x = "Axis.1", y = "Axis.2", color = "location", palette = "aaas",ellipse = TRUE, ellipse.level=.5,mean.point = F, mean.point.size = 5, star.plot = F) +labs(x = "PCoA 1 (24.1%) ", y = "PCoA 2 (20.6%)", colour = "location", fill = "location")+ annotate("text", x = -0.12, y = -0.2, label = label_perm, colour = "black")
pmain_brayF
```

##Arranging both diversity plots into one
```{r}
beta_div_plots = ggarrange(pmain_brayF,pmain_wuF,nrow = 2, ncol = 1, align="hv", common.legend = TRUE, labels = c("C","D")) 
beta_div_plots
```

#Diversity plot
Alpha and beta diversity plots arranged in one figure
```{r fig.width=10, fig.height=7}
diversity = ggarrange(alpha_div_plots, beta_div_plots, nrow = 1, ncol = 2, widths = c(1,1.5))
diversity
```

#Saving diversity plot as png format
```{r}
ggsave(plot = diversity, "~/OneDrive - Michigan State University/Zhang_lab/Marine_Iguana/figures-plots/microbiome/16S_diversity_iguana.png", width = 10.5, height = 7)
```


#COMPOSITION
##Phylum data frame
Clustering the abundance by sample at the Phylum level and transforming to relative abundance
```{r}
glom_phylum <- tax_glom(physeq, taxrank = 'Phylum')
data_phylum <- psmelt(glom_phylum) # create dataframe from phyloseq object

library(dplyr)
data_phylum_rel = data_phylum %>% group_by(Sample) %>% mutate(relAbundBySample = Abundance / sum(Abundance)*100) #Converting counts to relative abundance %
```

###Phylum Barplot
```{r}
#order of samples (otherwise it's based on the most abundant phyla per sample)
sample_names <- sort(unique(data_phylum$Sample))

phylum_ab <- ggbarplot(data=data_phylum_rel, x= "Sample", y = "relAbundBySample", color = "Phylum", fill = "Phylum", palette = "simpsons", ylab = "Relative Abundance", order = sample_names) +labs(colour = "Phylum", fill = "Phylum") + facet_wrap(~location, scales = "free_x") + theme(axis.text.x = element_text(angle = 90))

phylum_ab
```

###Phylum proportions
```{r}
phylum_proportions <- data_phylum_rel %>% 
  select(Phylum, relAbundBySample) %>% 
  group_by(Phylum) %>% 
  summarise(Proportion = mean(relAbundBySample)) 
phylum_proportions
```

##Family data frame
```{r}
glom_family <- tax_glom(physeq, taxrank = 'Family')
glom_family_rel = transform_sample_counts(glom_family, function(x) x/sum(x))
top20 <- names(sort(taxa_sums(glom_family), decreasing=TRUE)[1:20])
top20 #shows 20 results
family_top20 = prune_taxa(top20, glom_family_rel)
data_family <- psmelt(family_top20) # create dataframe from phyloseq object
```

###Family barplot
```{r fig.width=20, fig.height=5}
sample_names <- sort(unique(data_family$Sample))

family_ab <- ggbarplot(data=data_family, x= "Sample", y = "Abundance", color = "Family", fill = "Family", palette = get_palette("rickandmorty",20), ylab = "Relative Abundance", order = sample_names) +labs(colour = "Family", fill = "Family", position = position_fill()) + facet_grid(~location, scales = "free_x", space = "free") + theme(axis.text.x = element_text(angle = 90))

family_ab
```

###Family proportions
```{r}
family_proportions <- data_family %>% 
  select(Family, Abundance) %>% 
  group_by(Family) %>% 
  summarise(Proportion = mean(Abundance)) 
```

###Genus data frame
```{r}
glom_genus <- tax_glom(physeq, taxrank = 'Genus')
top20 <- names(sort(taxa_sums(glom_genus), decreasing=TRUE)[1:20])
top20 #shows 20 results

glom_genus_rel = transform_sample_counts(glom_genus, function(x) x/sum(x))

genus_top20 = prune_taxa(top20, glom_genus_rel)
data_genus <- psmelt(genus_top20) # create dataframe from phyloseq object
```

###Genus barplot
```{r fig.width=20, fig.height=5}
sample_names <- sort(unique(data_genus$Sample))

genus_ab <- ggbarplot(data=data_genus, x= "Sample", y = "Abundance", color = "Genus", fill = "Genus", palette = get_palette("simpsons",21), ylab = "Relative Abundance", order = sample_names,, position = position_stack()) +labs(colour = "Genus", fill = "Genus") + facet_grid(~location, scales = "free_x", space = "free") + theme(axis.text.x = element_text(angle = 90)) +  theme(legend.text = element_text(face = "italic"))

genus_ab
```

###Genus proportion
```{r}
genus_proportions <- data_genus %>% 
  select(Genus, Abundance) %>% 
  group_by(Genus) %>% 
  summarise(Proportion = mean(Abundance)) 
```

##Final composition plot with Family and genus relative abundance
```{r fig.width=20, fig.height=10}
composition_barplots <- ggarrange(family_ab,genus_ab, nrow = 2, labels = c("A","B"))
composition_barplots
```

##Saving composition plot as png
```{r}
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/figures-plots/microbiome")
ggsave("composition_marine-iguana_qiime.png", plot = composition_barplots, width = 20, height = 10)
```


#Saving composition plots by separate
```{r}
library(ggplot2)
setwd("/Users/karlavasco/OneDrive\ -\ Michigan\ State\ University/Zhang_lab/Marine_Iguana/figures-plots/microbiome")
ggsave("Genus_marine-iguana_qiime.png", plot = genus_ab, width = 15, height = 5)
ggsave("Family_marine-iguana_qiime.png", plot = family_ab, width = 15, height = 5)
ggsave("Phylum_marine-iguana_qiime.png", plot = phylum_ab, width = 15, height = 5)
```